{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-09-25 thesis/overview.tex
write more glosses

This section provides a convenient reference of terminology used throughout the thesis.
Use of {\em italics} within definitions will indicate terms also defined in this list.

@@

@@

		Short for {\em application programming interface} (which nobody ever says).

@@

@@

	\item {\bf Conflict.}

@@

@@

@@

@@

@@

		are both heuristic (\sect{\ref{sec:landslide-blocking}}),

@@

@@

@@

@@

@@

		The heuristic search ordering strategy used by {\em Quicksand}.

@@

@@

	\item {\bf Memory conflict.} See {\em conflict}.

@@

		Short for {\em mutual exclusion lock}. The name {\em P2} and {\em pthreads} use for their {\em lock API}.

@@

@@

@@

@@

@@

@@

@@

		A confusing term which, while CMU 15-410 teaches it to mean {\em concurrency bug},
		in research circles is generally taken to mean {\em data race}.

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

This section provides a convenient reference of terminology used throughout the thesis.
Use of {\em italics} within definitions will indicate terms also defined elsewhere in the list.

@@

	\item {\bf Actor.}
		(Also ``actor concurrency model''.)
		An independent execution component of a concurrent system.
		In {\em pthread}-like multithreaded programs, each {\em thread} is an actor.
		In distributed systems, each machine is an actor.

@@

		The set of function signatures which comprise the interface to a body of code.
		For example, the standard {\em mutex} API consists of
		{\tt mutex\_init()}, {\tt mutex\_lock()}, {\tt mutex\_unlock()},
		and optionally {\tt mutex\_trylock()} and {\tt mutex\_destroy()}.
		From ``application programming interface'' (which nobody ever says).

@@

	\item {\bf Completeness.}
		%Dual of {\em soundness}.
		In logic, said of an inference system in which any proposition consistent with the axioms can be proved.
		When speaking of {\em state space reduction},
		could mean opposite things depending on perspective:
		complete from an interleaving coverage perspective would mean not skipping any buggy interleavings;
		complete from a state space reduction perspective would mean not testing any two equivalent interleavings
		without pruning one.
		I define {\em soundness} as the dual from the reduction perspective (i.e., not pruning anything unsafe to prune),
		which is actually identical to the coverage perspective definition of completeness.
		To avoid confusion, this thesis avoids this term altogether,
		using {\em optimality} to refer to the reduction perspective definition instead.

@@

	\item {\bf Conflict.} See {\em memory conflict}.

@@

	\item {\bf Deadlock.}
		An execution state among a set of N {\em threads},
		where each thread is blocked waiting for an event that can be caused only by another thread in the set
		\cite{deadlock}.
		May arise either deterministically or concurrently; {\em Landslide} considers it a bug in any case.

@@

	\item {\bf Disjoint.}
		Said of a pair or of two sets of memory accesses
		when there are no two elements that share an address and also are not both reads.

@@

	\item {\bf Dynamic analysis.}
		Program analysis technique which involves analyzing concrete execution(s) of a program,
		either through tracing (via compiler-inserted instrumentation, simulation, or other means)
		or through post-hoc analysis of event logs.
		Searches for concrete evidence of undesirable program behaviour,
		and is blind to what the source code looks like (or even what language it be written in).
		Opposite of {\em static analysis}.

@@

		From ``estimated time of arrival''.

@@

		are both {\em heuristic} (\sect{\ref{sec:landslide-blocking}}),

@@

		Relation between two events executed by different threads in a single execution trace
		which captures the notion of concurrent execution.
		Comes in two flavours, {\em pure happens-before} and {\em limited happens-before}.

@@

		A testing strategy informed by empirical evidence rather than theoretically-founded algorithms.
		For example, {\em ICB} is a heuristic way of prioritizing {\em interleavings} with fewer {\em preemptions},
		based on the observation that most concurrency bugs in practice can be exposed with as few as 2-3 preemptions.
		For example, {\em Landslide} heuristically judges the program to be stuck in an infinite loop
		when it has executed at least 4000 times longer than previously-observed ``safe'' executions,
		based on the observation that no student has ever written correct thread library or kernel code
		that can exhibit such extreme interleaving-dependent variance.
	\item {\bf ICB.} Short for ``Iterative Context Bounding''. See \sect{\ref{sec:landslide-icb}}.

@@

		Said of two {\em transitions} by different threads,
		when they share no {\em memory conflicts};
		i.e., they can have no effect on each other's behaviour even if reordered.

@@

		A sequence of {\em transitions} of many threads representing the complete or partial execution
		of a {\em test case}.

@@

		The {\em heuristic} search ordering strategy and framework for incorporating {\em data race preemption points}
		used by {\em Quicksand}.

@@

		Variant of {\em happens-before} that relates two events only when a blocking/signalling synchronization affecting
		the second thread's runnability, such as {\tt cond\_wait()}/{\tt cond\_signal()} or {\tt thr\_fork()}
		occurs in between.
		In contrast with {\em pure happens-before},
		indicates not just that the events were observed
		to happen non-concurrently in the execution in question,
		but also that they could not be reordered even in alternate interleavings.

@@

	\item {\bf Memory conflict.}
		A pair of memory accesses between two {\em threads}, where at least one is a write,
		meaning that executing them in a different order from the one observed may produce different program behaviour.
		If both are writes, reordering them may cause a different value to be globally visible in that location
		to subsequent parts of the program.
		If one is a read and one is a write, reordering them may cause the read access to return a different value,
		possibly affecting that thread's subsequent logic.

@@

		Short for ``mutual exclusion lock''. The name {\em P2} and {\em pthreads} use for their {\em lock API}.

@@

	\item {\bf Optimality.} See {\em completeness}.

@@

		Operating systems class project in which students must implement synchronization primitives
		and thread lifecycle functions.

@@

		See \sect{\ref{sec:pebbles}}.

@@

	\item {\bf Preemption.}
		The act of interrupting one thread's execution to force a context switch to another.
		Can be performed either deliberately by the {\em model checker}
		or randomly by the {\em scheduler} as driven by system interrupts.
	\item {\bf Preemption point.}
		A single instance of a program state or code location where a {\em preemption point predicate} returns true.
		This thesis will also use this as shorthand for preemption point predicates,
		using the longer term only when the precise difference is important.
	\item {\bf Preemption point predicate.}

@@

		Variant of {\em happens-before} that relates two events when any synchronization operation introducing a memory
		barrier, which would prevent compiler or hardware access reorderings, occurs in between.
		In contrast with {\em limited happens-before},
		indicates that the events were observed to execute non-concurrently in this execution,
		but could possibly be reordered to be concurrent in alternate interleavings.

@@

		Testing framework which manages the execution of multiple {\em Landslide} instances,
		each configured with a different set of {\em preemption points},
		using the {\em Iterative Deepening} algorithm.
		Also supports various other execution modes in which only one Landslide instance is run at a time.
		See Chapter~\ref{chap:quicksand}.
		In later chapters, ``Landslide'' will refer to the project as a whole, including Quicksand,
		while the term ``Quicksand'' will be used only when specific reference to Iterative Deepening is required.

@@

		A confusing term which CMU 15-410 teaches to mean {\em concurrency bug},
		but in research circles is generally taken to mean {\em data race}.

@@

	\item {\bf Reduction.}
		The act of making a {\em state space} smaller by finding multiple execution sequences that can be proved equivalent,
		such that all but one may be skipped over.

@@

	\item {\bf Scheduler.}
		The component of an operating system kernel responsible for tracking the set of runnable threads,
		handling system interrupts,
		and allocating execution time among them (in round-robin, priority-driven, or other manner).
		May also to refer to the analogous component of {\em Landslide}
		which overrides the decision of the kernel scheduler.
	\item {\bf Soundness.}
		Said of a search algorithm
		which provides the property that if a bug exists under the given test case,
		given arbitrary CPU time,
		it will eventually be found;
		in other words, admits no {\em false negatives}.
		When said of a {\em reduction} strategy as applies to an existing search algorithm,
		means that any bugs found by the original algorithm will still be found under the reduction,
		even if there were false negatives
		(e.g., applying {\em DPOR} without using {\em data race preemption points}).
		Used by analogy with logic, in which said of an inference system,
		means contradiction cannot be derived from the axioms.
		See also {\em completeness}.

@@

		Concurrency testing {\em dynamic analysis} technique involving controlling a system's nondeterminism
		to exhaustively check all possible thread interleavings.
		See \sect{\ref{sec:overview-stateless}}.
		Often ``MC'' for short in this thesis (to mean either ``model checking'' or ''model checker'' depending on context).
		This name is potentially confusing for tools which explore {\em state spaces}:
		the term ``stateless'' refers to the implicit way they use {\em DPOR} to ensure {\em sound} checking
		of all possible program states, without actually explicitly storing those states in memory.
		The second half of the name is another historical accident,
		as most modern MCs do not check ``models'' in the sense that
		programs be verified against external formal specifications,
		but rather the program's internal assertions serve as informal specifications,
		alongside the MC's own bug-detection predicates.

@@

	\item {\bf Static analysis.}
		Class of program analysis techniques which process the source code,
		usually at syntax-tree or LLVM IR level,
		looking for suspicious implementation patterns,
		rather than actually observing behaviours that arise during execution.
		Opposite of {\em dynamic analysis}.

@@

		A student who volunteers to help run a class, holding office hours, grading projects, and so on,
		compensated either with course credit or pay from the university.
		Probably has taken the class herself in the past.

@@

		A piece of code, separate from the code under test,
		to drive execution of the code under test according to a certain pattern designed to expose interesting behaviour.
		In the context of {\em P2}, the thread library is the code under test,
		and client code thereof, such as {\em paradise\_lost.c}, is the test case.

@@

		Unit of concurrent execution in a typical modern program,
		concretely represented as a set of private CPU registers, including program counter.
		Probably has its own stack as well, allocated by the userspace thread library.
		A program may have as many parts of its code executing simultaneously as threads exist.

@@

		Data structure for time-keeping in distributed systems which uses a list of epoch numbers
		to track when the execution of any given pair of {\em actors} should be considered simultaneous/concurrent.
		Used in {\em Landslide} to implement {\em pure happens-before}.
